---
title: "NEB competition code"
author: "Sam Siljee"
date: "2024-05-14"
output: html_document
---

## NEB competiton code

This R markdown document contains the code to produce the 2 minute video clip used for the NEB passion in science competition entry.

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Libraries
library(tuneR)
library(mzR)
library(dplyr)
library(ggplot2)
library(av)
library(data.table)

# Source functions
source("functions.R")

# Set some variables
sample_rate <- 44100
duration <- 1
start_time <- 1528.560
end_time <- 1645.594
```

## Load and process the data
First I need to trim out the 2 minutes of interest from the whole 90 minute test run. The time limit on the submitted video is two minutes, and it will significantly speed up computation.
Retention time of clip start: 25 min ?

```{r Load and process the data, inclue = FALSE}
# Open test MS file
ms_data <- openMSfile("test.mzML")

# Extract all the peaks
ms_peaks <- peaks(ms_data)

# Extract corresponding header information
ms_header <- header(ms_data) %>%
  filter(retentionTime > start_time &
    retentionTime < end_time)

# Get list of MS1 and MS2 spectrum indexes, and filter for rention times in the 2 minute segment
ms_1_indexes <- filter(ms_header, msLevel == 1) %>% .$seqNum
ms_2_indexes <- filter(ms_header, msLevel == 2) %>% .$seqNum
```

## Generate the waveforms
Next the waveforms should be generate for each spectrum, and saved as .txt files.

```{r Generate waveforms, include = FALSE}
# Synthesise waveforms, and write to .txt (so big runs can be interrupted)
for (i in ms_header$seqNum) {
  write(
    spectrum_to_waveform(ms_peaks[[i]], duration = duration, sampling_rate = sample_rate),
    ncolumns = 1,
    file = paste0("NEB_competition_waveforms/waveform_", i, ".txt")
  )
}
```

## Generate soundscape
Once the waveforms have been saved, they can be added together to form the soundscape.

```{r Generate soundscape, include = FALSE}
# Total time of the clip, at set sample rate
max_RT <- max(ms_header$retentionTime)
min_RT <- min(ms_header$retentionTime)
total_time <- round(sample_rate * (max_RT - min_RT + 1))

# Make a stereo compilation, Left channel for MS1 spectra, Right channel for MS2 spectra
# Initialise blank waveform
ms1_waveform <- rep(0, total_time + duration * sample_rate)

# Read in the .txt files and add to the waveform
for (i in ms_1_indexes) {
  # Generate blank time before and after the waveform
  RT_pre <- rep(0, round(sample_rate * (ms_header$retentionTime[ms_header$seqNum == i] - min_RT)))
  RT_post <- rep(0, total_time - length(RT_pre))

  # read in the waveform and account for intensity
  waveform <- read.delim(paste0("NEB_competition_waveforms/waveform_", i, ".txt"), header = FALSE)[, 1] * ms_header$totIonCurrent[ms_header$seqNum == i]

  # Add to the waveform with blank RT sound
  ms1_waveform <- ms1_waveform + c(RT_pre, waveform, RT_post)

  # Monitor progress
  print(i)
}

# Normalise the waveform
ms1_waveform <- (ms1_waveform / max(abs(ms1_waveform))) * 32000

# Write the MS1 compiled waveform to disk
write(
  ms1_waveform,
  ncolumns = 1,
  file = "NEB_competition_waveforms/ms1_compiled_waveforms_.txt"
)

# Repeat for MS2 scans
# Initialise blank waveform
ms2_waveform <- rep(0, total_time + duration * sample_rate)

# Read in the .txt files
for (i in ms_2_indexes) {
  # Generate blank time before and after the waveform
  RT_pre <- rep(0, round(sample_rate * (ms_header$retentionTime[ms_header$seqNum == i] - min_RT)))
  RT_post <- rep(0, total_time - length(RT_pre))

  # read in the waveform and account for intensity
  waveform <- read.delim(paste0("NEB_competition_waveforms/waveform_", i, ".txt"), header = FALSE)[, 1] * ms_header$totIonCurrent[ms_header$seqNum == i]

  # Add to the waveform with blank RT sound
  ms2_waveform <- ms2_waveform + c(RT_pre, waveform, RT_post)

  # Monitor progress
  print(i)
}

# Normalise the waveform
ms2_waveform <- (ms2_waveform / max(abs(ms2_waveform))) * 32000

# Write the MS2 compiled waveform to disk
write(
  ms2_waveform,
  ncolumns = 1,
  file = "NEB_competition_waveforms/ms2_compiled_waveforms_.txt"
)

# Create stereo wav object
stereo_wave_obj <- Wave(
  left = round(ms1_waveform),
  right = round(ms2_waveform),
  samp.rate = sample_rate,
  bit = 16
)

# Save as .wav
writeWave(stereo_wave_obj, file = "NEB_competition_2_minute_audio.wav")
```

## Visualisation
This section is where the frames for the visualisation will be made

```{r Visualisation plot method, include = FALSE}
# For MS1 scans, plot a single line, otherwise plot the spectra back-to-back, MS1 on the left, MS2 on the right
# Set a value to scale the contrast vector by
max_contrast <- log(max(ms_header$totIonCurrent))

for (i in ms_header$seqNum) {
  # Plot with a blank right side if MS1 spectrum
  if (ms_header$msLevel[ms_header$seqNum == i] == 1) {
    double_plot(
      spectrum_left = as.data.frame(ms_peaks[[i]]),
      spectrum_right = data.frame(mz = min(ms_peaks[[i]][, 1]), intensity = 1),
      contrast = log(ms_header$totIonCurrent[ms_header$seqNum == i]) / max_contrast
    )
    # Otherwise plot the corresponding MS1 spectrum to the left, and the MS2 plot to the right
  } else {
    double_plot(
      spectrum_left = as.data.frame(ms_peaks[[ms_header$precursorScanNum[ms_header$seqNum == i]]]),
      spectrum_right = as.data.frame(ms_peaks[[i]]),
      precursor_mz = ms_header$precursorMZ[ms_header$seqNum == i],
      precursor_intensity = ms_header$precursorIntensity[ms_header$seqNum == i],
      contrast = log(ms_header$totIonCurrent[ms_header$seqNum == i]) / max_contrast
    )
  }

  # Save the plots
  ggsave(paste0("NEB_competition_plots/Plot_", i, ".png"), width = 1920, height = 1080, units = "px")
}

# Put the plots together in video format
png_files <- paste0("NEB_competition_plots/Plot_", ms_header$seqNum, ".png")
av_encode_video(
  png_files,
  "NEB_competition_video.mp4",
  framerate = length(ms_header$seqNum) / (end_time - start_time),
  audio = "NEB_competition_2_minute_audio.wav"
)
```

This seems to produce a servicable video, although one thing lacking is a strong connection between the audio and the video. I suspect that the timing will be somewhat off, ideally I'd use the retention times to trigger the video frame - but I'm not sure if this is possible with the `av` package as is.
Another disconnect is the scaling of the plots - there is no reflection of an increase in loudness, perhaps I should scale over the whole video rather than each individual plot.

I was not completely satisfied with the contrast effect, so I manually scaled and trimmed the contrast vector to get slightly more range.
```{r Visualisation plot method 2 scaled contrast, include = FALSE}
# For MS1 scans, plot a single line, otherwise plot the spectra back-to-back, MS1 on the left, MS2 on the right

# Create a vector of contrasts
con_v <- log(ms_header$totIonCurrent)

# Trim out the top and bottom 10% of spectra to spread more evenly
con_max <- sort(con_v)[length(con_v) * 0.9]
con_min <- sort(con_v)[length(con_v) * 0.1]
con_v[con_v > con_max] <- con_max
con_v[con_v < con_min] <- con_min

# Scale from 0.01 to 1
con_v <- (con_v - min(con_v)) / max(con_v - min(con_v)) * (1 - 0.01) + 0.01
# Assign to a vector that matches the `ms_header$seqNum` vector
contrast_vector <- NULL
index <- 1
for (i in ms_header$seqNum) {
  contrast_vector[i] <- con_v[index]
  index <- index + 1
}

for (i in ms_header$seqNum) {
  # Plot with a blank right side if MS1 spectrum
  if (ms_header$msLevel[ms_header$seqNum == i] == 1) {
    double_plot(
      spectrum_left = as.data.frame(ms_peaks[[i]]),
      spectrum_right = data.frame(mz = min(ms_peaks[[i]][, 1]), intensity = 1),
      contrast = contrast_vector[i]
    )
    # Otherwise plot the corresponding MS1 spectrum to the left, and the MS2 plot to the right
  } else {
    double_plot(
      spectrum_left = as.data.frame(ms_peaks[[ms_header$precursorScanNum[ms_header$seqNum == i]]]),
      spectrum_right = as.data.frame(ms_peaks[[i]]),
      precursor_mz = ms_header$precursorMZ[ms_header$seqNum == i],
      precursor_intensity = ms_header$precursorIntensity[ms_header$seqNum == i],
      contrast = contrast_vector[i]
    )
  }

  # Save the plots
  ggsave(paste0("NEB_competition_plots_contrast/Plot_", i, ".png"), width = 1920, height = 1080, units = "px")
}

# Put the plots together in video format
png_files <- paste0("NEB_competition_plots_contrast/Plot_", ms_header$seqNum, ".png")
av_encode_video(
  png_files,
  "NEB_competition_video_scaled_contrast_plots_2.mp4",
  framerate = length(ms_header$seqNum) / (end_time - start_time),
  audio = "NEB_competition_2_minute_audio.wav"
)
```


## Image as visualistaion
This method plots the spectra as crossing lines of various intensity as an alternative to back-to-back plotting

```{r Visualisation - image method, include = FALSE}
for (i in ms_header$seqNum[ms_header$seqNum > 7950]) {
  # Open png device
  png(paste0("NEB_competition_images/Image_", i, ".png"), width = 1920, height = 1080)
  # If MS1, create an image with MS1 for both
  if (ms_header$msLevel[ms_header$seqNum == i] == 1) {
    double_image(
      spectrum_1 = as.data.frame(ms_peaks[[i]]),
      spectrum_2 = as.data.frame(ms_peaks[[i]])
    )
  } else {
    # Image the spectrum with MS1 along the y axis, and MS2 on the x axis
    double_image(
      spectrum_1 = as.data.frame(ms_peaks[[ms_header$precursorScanNum[ms_header$seqNum == i]]]),
      spectrum_2 = as.data.frame(ms_peaks[[i]])
    )
  }
  # Close the device
  dev.off()
}

# Put the plots together in video format
png_files <- paste0("NEB_competition_images/Image_", ms_header$seqNum, ".png")
av_encode_video(
  png_files,
  "NEB_competition_video_image_method.mp4",
  framerate = length(ms_header$seqNum) / (end_time - start_time),
  audio = "NEB_competition_2_minute_audio.wav"
)
```

## Thumbnail image
Generating the thumbnail image to go with the entry

```{r Thumbnail image, include = FALSE}
double_plot(
  spectrum_left = as.data.frame(ms_peaks[[ms_header$seqNum[1]]]),
  spectrum_right = as.data.frame(ms_peaks[[ms_header$seqNum[2]]]),
  precursor_mz = ms_header$precursorMZ[ms_header$seqNum == ms_header$seqNum[2]],
  precursor_intensity = ms_header$precursorIntensity[ms_header$seqNum == ms_header$seqNum[2]],
  contrast = 0.7
)

# Save the plots
  ggsave(paste0("NEB_competition_thumbnail_image.jpg"), width = 10000, height = 10000, units = "px")
```

