---
title: "NEB competition code"
author: "Sam Siljee"
date: "2024-05-14"
output: html_document
---

## NEB competiton code

This R markdown document contains the code to produce the 2 minute video clip used for the NEB passion in science competition entry.

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Libraries
library(tuneR)
library(mzR)
library(dplyr)
library(ggplot2)

# Source functions
source("functions.R")
```

## Subset data
First I need to trim out the 2 minutes of interest from the whole 90 minute test run. The time limit on the submitted video is two minutes, and it will significantly speed up computation.

```{r Subset data, inclue = FALSE}

```

## Load data
After the data has been subset and saved as an RDS file, it can be loaded more quickly for future work.

```{r Load data, include = FALSE}

```

## Generate the waveforms
Next the waveforms should be generate for each spectrum, and saved as .txt files.

```{r Generate waveforms, include = FALSE}

```

## Generate soundscape
Once the waveforms have been saved, they can be added together to form the soundscape.

```{r Generate soundscape, include = FALSE}

```