---
title: "The Sound of Science"
author: "Sam Siljee"
date: 9 Oct 2024
format:
  revealjs:
    transition: slide
  # pptx
editor: visual
---

```{r, echo = FALSE}
library(tidyverse)
source("functions.R")
spectrum <- read.csv("test_spectra/spectrum_1.csv")
theme_set(theme_void())
```

## Introduction to Mass Spectrometry

```{r}
spectrum %>%
  ggplot(aes(x = mz, y = intensity)) +
  geom_col() +
  theme_bw()
```
But there's a third dimension...

## Coersion
- Tone
- Pitch
- Quantisation

. . .

- Honour the original data

## Sonification process

Fourier theorem: any complex waveform can be made up from the addition of simpler waveforms of different frequencies

. . .

```{r}
#| width: "100%"

data.frame(
  x = 1:1000,
  y_1 = sin(0.1 * 1:1000),
  y_2 = cos(0.02 * 1:1000)
) %>%
  ggplot() +
  geom_line(aes(x = x, y = y_1)) +
  geom_line(aes(x = x, y = y_2 + 1*2))
```

##

```{r waves 2}
data.frame(
  x = 1:1000,
  y_1 = sin(0.1 * 1:1000),
  y_2 = cos(0.02 * 1:1000)
) %>%
  mutate(y_3 = (y_1 + y_2) / max(abs(y_1 + y_2))) %>%
  ggplot() +
  geom_line(aes(x = x, y = y_1)) +
  geom_line(aes(x = x, y = y_2 + 1*2)) +
  geom_line(aes(x = x, y = y_3 + 2*2), col = "darkgreen") +
  theme_void()
```

##

```{r waves 3}
data.frame(
  x = 1:1000,
  y_1 = sin(0.2 * 1:1000),
  y_2 = sin(0.4 * 1:1000) * 0.6,
  y_3 = sin(0.8 * 1:1000) * 1.3,
  y_4 = sin(1.6 * 1:1000) * 0.3,
  y_5 = sin(3.2 * 1:1000)
) %>%
  mutate(y_6 = (y_1+y_2+y_3+y_4+y_5) / max(abs(y_1+y_2+y_3+y_4+y_5))) %>%
  ggplot() +
  geom_line(aes(x = x, y = y_1)) +
  geom_line(aes(x = x, y = y_2 + 1*2)) +
  geom_line(aes(x = x, y = y_3 + 2*2)) +
  geom_line(aes(x = x, y = y_4 + 3*2)) +
  geom_line(aes(x = x, y = y_5 + 4*2)) +
  geom_line(aes(x = x, y = y_6 + 5*2), col = "darkgreen")
```

##

```{r spectrum repeat}
spectrum %>%
  ggplot(aes(x = mz, y = intensity)) +
  geom_col() +
  theme_bw()
```

##

```{r spectrum waveform}
data.frame(
  x = 1:10000,
  y = spectrum_to_waveform(spectrum)[34101:44100]
) %>%
  ggplot() +
  geom_line(aes(x = x, y = y))
```

## Complexity
- Emergent properties

## Interpretation
- AI and directly engaging with the data
- Methodology vs biology

## Practical Implications
- Learning about MS
- Vision impaired
- Accessibility
- While this won't replace conventional analysis any time soon, our ability to understand data is outstripped by our ability to generate data. New ideas have to start somewhere

## Where to Next
- Develop methodology further - reintegrating processed data (bringing the biology back)
- Present at ACMC and Leonardo
- Working with artists

## Acknowledgements
- Gillies McIndoe Research Institute
- Te Herenga Waka - Victoria University of Wellington
- Wellington Doctoral Scholarship
