---
title: "The Sound of Science"
author: "Sam Siljee"
date: 9 Oct 2024
format:
  revealjs:
    transition: slide
  # pptx
editor: visual
---

```{r}
library(tidyverse)
source("functions.R")
spectrum <- read.csv("test_spectra/spectrum_1.csv")
theme_set(theme_void())
```

## Introduction to Mass Spectrometry

```{r}
spectrum %>%
  ggplot(aes(x = mz, y = intensity)) +
  geom_col() +
  theme_bw()
```

## Sonification process

As scientists we are always curious to understand how something works.

Here is the magic

. . .

```{r}
#| width: "100%"

data.frame(
  x = 1:1000,
  y_1 = sin(0.1 * 1:1000),
  y_2 = cos(0.02 * 1:1000)
) %>%
  ggplot() +
  geom_line(aes(x = x, y = y_1)) +
  geom_line(aes(x = x, y = y_2 + 1*2), col = "blue")
```

## Sonification process

```{r waves 2}
data.frame(
  x = 1:1000,
  y_1 = sin(0.1 * 1:1000),
  y_2 = cos(0.02 * 1:1000)
) %>%
  mutate(y_3 = (y_1 + y_2) / max(abs(y_1 + y_2))) %>%
  ggplot() +
  geom_line(aes(x = x, y = y_1)) +
  geom_line(aes(x = x, y = y_2 + 1*2), col = "blue") +
  geom_line(aes(x = x, y = y_3 + 2*2), col = "red") +
  theme_void()
```

## Sonification process

```{r waves 3}
data.frame(
  x = 1:1000,
  y_1 = sin(0.1 * 1:1000),
  y_2 = sin(0.07 * 1:1000) * 0.6,
  y_3 = sin(0.23 * 1:1000) * 1.3,
  y_4 = sin(0.234 * 1:1000) * 0.3,
  y_5 = sin(0.017 * 1:1000)
) %>%
  mutate(y_6 = (y_1+y_2+y_3+y_4+y_5) / max(abs(y_1+y_2+y_3+y_4+y_5))) %>%
  ggplot() +
  geom_line(aes(x = x, y = y_1)) +
  geom_line(aes(x = x, y = y_2 + 1*2), col = "blue") +
  geom_line(aes(x = x, y = y_3 + 2*2), col = "orange") +
  geom_line(aes(x = x, y = y_4 + 3*2), col = "green") +
  geom_line(aes(x = x, y = y_5 + 4*2), col = "yellow") +
  geom_line(aes(x = x, y = y_6 + 5*2), col = "red")
```

## Sonification process

```{r spectrum repeat}
spectrum %>%
  ggplot(aes(x = mz, y = intensity)) +
  geom_col() +
  theme_bw()
```

## Sonification process

```{r spectrum waveform}
data.frame(
  x = 1:10000,
  y = spectrum_to_waveform(spectrum)[34101:44100]
) %>%
  ggplot() +
  geom_line(aes(x = x, y = y))
```

## Assumptions

## Complexity

## Interpretation

-   AI and directly engaging with the data

## Practical Implications

## Where to Next

-   Develop methodology further - reintegrating processed data (bringing the biology back)
-   Present at ACMC and Leonardo
-   Working with artists

## Acknowledgements
